---
title: "Homework 4 (Final Project)"
author: "Due date: 12/19/25"
output:
  html_document:
    code_folding: show
    toc: true
    toc_float: true
    theme: "flatly"
editor_options:
  chunk_output_type: console
---

# Homework 4 — Final Project: End-to-End Analysis of a Dataset of Your Choice

For this final project, you will choose a dataset of personal or research interest and demonstrate mastery of the analytical and machine-learning techniques introduced in this course.  
Your work should be submitted as an **R Markdown (`.Rmd`) file** with rendered **HTML output**, including narrative explanations, code, and figures.

---

## **Project Overview & Requirements (100 points)**

### **1. Dataset Description and Research Questions (20 pts)**

Select **any dataset of interest**—from your own research, publicly available repositories, or other sources.  

Provide the following:

- A **clear description of the dataset**, including:
  - Number of observations  
  - Variables/features included  
  - The scientific or practical context  
- A **clearly defined hypothesis or set of questions** you aim to answer  

**Write at least one paragraph** describing:
- Why this dataset interests you  
- What specific relationships, predictions, or structure you intend to investigate  

```{r data-description, eval=FALSE}
# Example structure (students replace with their own)
# mydata <- read_csv("mydata.csv")
# glimpse(mydata)
```

### **2. Apply at Least Four Machine Learning Approaches (40 pts)**
You must apply **at least four techniques** covered in the class:

* **At least 1 unsupervised method** (e.g., PCA, t-SNE, UMAP, clustering, dimension reduction)
* **At least 1 supervised method** (e.g., SVM, Random Forest, XGBoost, Logistic Regression, Neural Networks)
* The remaining methods may be either supervised or unsupervised.

For **each** technique, include:

* A short explanation of why you chose it
* The R code implementing the method
* Visualization(s) illustrating the method
* A brief interpretation of your results

**Example section template** (students should copy/modify this):

#### **2.X Method Name (e.g., PCA, SVM, RF, etc.)**
Explanation (what this method does and why you used it)

```{r}
# Your code here
```


### **3. Visualizations (20 pts)** 
Your project must include multiple high-quality visualizations, such as:

* PCA plots
* Heatmaps
* Variable importance plots
* ROC curves
* Confusion matrices
* Cluster dendrograms
* SHAP/LIME explanations
* Neural-network architecture or training curves

Each visualization should be accompanied by a caption or short explanation.

##### Visualization(s)
Introduce each of your plots

```{r}
#Plots here
```

Interpretation and briefly summarize what this method reveals about your dataset.

### **4. Summary & Conclusions (20 pts)** 
Provide a coherent summary of:

* Your dataset
* The machine learning methods you applied
* What each method contributed to your understanding
* Which techniques performed best and why
* Limitations and what future analysis might look like

This section should be **at least 2–3 paragraphs**.


### Deliverables
Please Submit:

1. Completed .Rmd file
2. Rendered HTML report
3. All code, narrative text, visualizations, and interpretations included

### Grading Summary

| Component                                                                 | Points |
|---------------------------------------------------------------------------|--------|
| Dataset description & hypotheses                                          | 20     |
| Application of ≥4 ML methods (≥1 unsupervised, ≥1 supervised)             | 40     |
| Plots & visualizations for each method                                    | 20     |
| Summary of results                                                        | 20     |
| **Total**                                                                 | **100**|
